The Galaxy Clustering Crisis in Abundance Matching

### 2017-05-02 paper/paper.tex
updates

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Nikhil Padmanabhan$^{2}$, and 
Frank C. van den Bosch$^{1}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
}

@@

\begin{abstract}
Galaxy clustering is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM methods based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, simple $V_{\rm peak}$ based SHAM models have their own problems, and there is little theoretical reason to prefer $V_{\rm peak}$ over $M_{\rm peak}$ as a proxy for stellar mass.  We identify three possible remedies that could save mass based models.  1.) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption of sub-haloes in modern dark matter simulations, 2.) satellites must grow significantly after the time of accretion, and/or 3.) stellar mass is significantly affected by halo assembly history.  Regardless of the particulars, popular SHAM models based on $M_{\rm peak}$ cannot be correct as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

SHAM, in its most simple form, assumes that the cumulative abundance of galaxies and haloes can be used to map galaxy properties uniquely onto (sub-)haloes.  Using stellar mass, $M_{*}$, and a halo mass proxy $\mathcal{M}$, SHAM assumes:
%
\begin{linenomath}
\begin{equation}
\label{eq:sham}
N(>M_*) = N(>\mathcal{M}) \implies M_{*} = f(\mathcal{M})
\end{equation} 
\end{linenomath}
%
where $f(\mathcal{M})$ is some monotonically increasing function.  this maps the most massive galaxies to the most massive haloes.  The function $f(\mathcal{M})$ may be determined non-parametrically by matching the rank orders of galaxies and haloes.  We refer to this method of SHAM as ``rank order'' abundance matching.  However, we also consider another class of SHAM models where $f(\mathcal{M})$ is parametrized.  This introduces the possibility that the observed abundance of galaxies strictly matched; however, it is generally the goal of such models to preserve this characteristic.  We consider these two types of models similar enough to be contained within the class of models we refer to as SHAM.

@@

\cite{Moster:2013ab} parametrize the SMHM relation as a function of halo mass and redshift to account for evolution.  Furthermore, they assume the stellar mass of satellite galaxies is determined by the mass of the sub-halo at the time of its accretion into a more massive host-halo for the first time.  The functional form is given by:
%
\begin{linenomath}
\begin{equation}
\label{eq:moster_smhm}
\frac{\langle M_{*} | \mathcal{M}\rangle(a)}{\mathcal{M}} = 2N\left[ \left(\frac{\mathcal{M}}{{M}_1} \right)^{-\beta} + \left( \frac{\mathcal{M}}{{M}_1} \right)^{\gamma} \right]^{-1}
\end{equation}
%
\end{linenomath}
%
\begin{linenomath}
\begin{align}
\label{eq:smhm_params}
\log[{M}_1(a)]& = {M}_{10} + {M}_{11}(1-a) \\
N(a) &= N_{10} + N_{11}(1-a) \\
\beta(a) &= \beta_{10} + \beta_{11}(1-a) \\ 
\gamma(a) &=  \gamma_{10} + \gamma_{11}(1-a)
\end{align}
\end{linenomath}
%
where the scale factor is either the instantaneous scale factor, or the one at the time of accretion:
%
\begin{linenomath}
\begin{equation}
a = 
\begin{cases}
    a      & \quad \text{if host-halo}\\
    a_{\rm acc}  & \quad \text{if sub-halo} \\
  \end{cases}
\end{equation}
\end{linenomath}
%
and $\mathcal{M}$ is the instantaneous mass or the mass at accretion, $M_{\rm acc}$, for sub-haloes.  Scatter in the SMHM relation is modelled as a fixed log-normal with $\sigma \approx 0.18$.

@@

\citet{Behroozi:2013fg} make similar assumptions as M13 but utilize a different parametrization given by:
%
\begin{linenomath}
\begin{equation}
\log[\langle M_*|\mathcal{M} \rangle(a)] = \log(\epsilon M_1) + f(\log(\mathcal{M}/M_1)) - f(0)
\end{equation}
\end{linenomath}
%
where,
% 
\begin{linenomath}
\begin{equation}
f(x) = -\log(10^{\alpha x}+1) + \delta\frac{\left[ \log(1+\exp[x])\right]^{\gamma}}{1+\exp(10^{-x})}
\end{equation}
\end{linenomath}
%
and where $\mathcal{M} = M_{\rm peak}$.  The parameters evolve with redshift as:
%
\begin{linenomath}
\begin{align}
\nu(a) & = e^{-4a^2}  \\
\log[M_1(a)] & =  M_{10} + \nu\left[ M_{1,a}(a-1) + M_{1,z}(z) \right] \\
\log[\epsilon(a)] & = \epsilon_0 + \nu \left[\epsilon_a(a-1) +\epsilon_z (z) \right] + \\ \nonumber 
& ~~~~ \epsilon_{a,2}(a-1)  \\
\alpha(a) & =  \alpha_{0} + \nu\left[ \alpha_a(a-1)\right] \\
\delta(a) & =   \delta_{0} + \nu\left[ \delta_a(a-1) + \delta_z (z)\right] \\
\gamma(a) &= \gamma_{0} + \nu\left[ \gamma_a(a-1) + \gamma_z (z)\right]
\end{align}
\end{linenomath}
% 
In fitting their model, \cite{Behroozi:2013fg} also allow the scatter in the SMHM relation to vary as a function of redshift.  The variation found is consistent with no variation (constant over cosmic time), so we neglect the small variation in the best fit model and use a constant non-varying scatter.  We have checked that including this variation has no appreciable effect on our conclusions.   As shorthand, we will refer to this model as ``B13''.   B13 constrained the parameters of this model by fitting the stellar mass function at multiple epochs, specific star-formation rates, and the cosmic star-formation history between z=0-8.  The values used in this paper are taken directly from \cite{Behroozi:2013fg} and listed in table 2.

@@

\cite{Yang:2012ew} take a different approach than M13 and B13.  They allow for evolution with redshift in a similar manner, but additionally allow satellites to grow or lose mass after the time of accretion.  The SMHM relation for central galaxies is: 
%
\begin{linenomath}
\begin{equation}
\langle M_{*,\rm cen} |  \mathcal{M} \rangle(z) = M_{0}\frac{\left( \mathcal{M}/M_1 \right)^{\alpha+\beta}}{\left(1+ \mathcal{M}/M_1 \right)^{\beta}}
\end{equation}  
\end{linenomath}
%
where $\mathcal{M}=M_{180b}$, and the parameters\footnote{We have altered the naming scheme somewhat to be more consistent with the other models.} evolve with redshift as:
%
\begin{linenomath}
\begin{align}
\log[M_{0}(z)] &= M_{01} + \gamma_1 z \\
\log[M_{1}(z)] &= M_{11} + \gamma_2 z \\
\alpha(z) &= \alpha_0 + \gamma_3 z \\
\log[\beta(z)] &= \log(\beta_0) + \gamma_4 z + \gamma_5 z^2
\end{align}
\end{linenomath}
%
The mean stellar mass of satellites is determined by interpolating between the stellar mass a satellite would have had at the time of accretion and the stellar mass of a central galaxy with the same halo mass as the satellite at accretion.  This value is determined by a single parameter, $c$:
%
\begin{linenomath}
\begin{align}
\label{eq:y12sat_model}
\langle M_{*,\rm sat}(z, z_{\rm acc})\rangle =&  (1-c)\langle M_{*,\rm cen} |  \mathcal{M}_{\rm acc} \rangle(z_{\rm acc})~+ \\ \nonumber &  c\langle M_{*,\rm cen}|  \mathcal{M} \rangle(z)
\end{align}
\end{linenomath}
%
where $\mathcal{M}_{\rm acc}$ is the halo mass proxy at the time of accretion for the sub-halo.  The case where $c=0$ corresponds to no growth (or mass loss) since the time of accretion.  The case where $c=1$ corresponds to using the same SMHM relation for centrals and satellites at all redshifts.

@@

Furthermore, each of the evolving models was tuned using observations based on different methods of determining stellar mass.  We apply a simple set of conversions in order to homogenize the stellar masses to a single system as described in Appendix \ref{appendix:stellar_mass_conv}.   We show the stellar mass function for a mock based on each model in Fig. \ref{fig:phi}.  The stellar mass functions in each mock vary by up to $\sim 0.2$ dex from our fiducial stellar mass function used to create the RM and RV mocks.  This is the result of each model being tuned with differing:
\begin{enumerate}
\item stellar mass functions, $\phi(M_*,z)$
\item halo mass definitions/(sub-)halo finder
\item cosmologies
\item simulations which are subject to cosmic variance
\end{enumerate}
While in principle each model could be re-fit using the same stellar mass function(s) and simulation, that is beyond the scope of this paper.  We check that simple alterations to the parameters of each model which bring each into better agreement with the LW09 stellar mass function do not have a significant effect on the clustering signal predictions presented in \S \ref{sec:clustering}.  Given this, and the complexity involved in re-fitting each model, we use the parameters as they are quoted in each paper (and reproduced in table \ref{table:model_parameters}), with only minor alterations as discussed in the sections above.

@@

Here we compare the RV model to the other models by examining the relation between stellar mass and halo mass that arises in each model, in the RV model by correlation between $V_{\rm peak}$ and $M_{\rm peak}$, and in the other models by design.  In the top panel of Fig. \ref{fig:SMHM_comparison}, we examine the relation between stellar mass and $M_{\rm peak}$ for individual galaxies in the RV mock.  Beyond the ubiquitous shape of the SMHM relation, the most obvious additional trend is that the RV model predicts that sub-haloes ($z_{\rm acc}>0$) host more massive galaxies than than host-haloes of equivalent $M_{\rm peak}$.  This is precisely the expectation given that sub-haloes generally have larger values of $V_{\rm peak}$ than host-haloes as shown in the previous section.

@@

We calculate the sub-halo completeness in the simulation as the ratio of the empirical $f_{\rm sub}$ and the fit using eq. \ref{eq:fsub_fit} as shwon in the bottom panel of Fig. \ref{fig:f_sub}.  We then model the completeness, $C_{\rm sub}$, as a function of halo mass as:
%
\begin{linenomath}
\begin{equation}
\label{eq:sub_completeness}
C_{\rm sub}(M_{\rm peak}) = \frac{C_0}{1.0 + \left(\frac{M_0}{M_{\rm peak}}\right)^{\gamma}}
\end{equation}
\end{linenomath}
%
We find $\log(M_0/h^{-1}M_{\odot})=9.980\pm0.003$ and $\gamma=1.27 \pm 0.02$ provide a good fit, as shown by the red line in the bottom panel in Fig. \ref{fig:f_sub}.  For now, we fix $C_0=1.0$, and we examine the possibility of $C_0<1.0$ at the end of this section.  The implicit assumption here is that when sub-haloes are well resolved, as is the case for massive (sub-haloes) at the time of infall, there should be no missing sub-structure.  Our estimation of the completeness is broadly consistent with the 1000 particle threshold found by \citet{Guo:2013fm} using the Millennium simulation suite.  Furthermore, we find that $C_{\rm sub}(M_{\rm peak})$ is very nearly constant with redshift.

@@

The Y12 model model when fit to galaxy clustering in \cite{Yang:2012ew}, which used an analytic halo model and let the sub-halo abundance be a free parameter, finds results which are consistent with our finding here that $C_0 \sim 0.5$, namely that the sub-halo abundance needs to be approximately a factor of two larger than that in Bolshoi.  Furthermore, \cite{Jiang:2016dw} find that the unevolved (peak sub-halo mass) surviving (not disrupted) sub-halo mas function is approximately a factor of two lower than the unevolved sub-halo mas function.  This suggest a plausible source of the missing sub-haloes needed to host orphans could be artificially disrupted sub-haloes.

@@

Assembly bias is a phenomenon observed in dark-matter simulations that the clustering of haloes is effected by formation history in addition to mass \citep{Gao:2005ds, Wechsler:2006rg, Gao:2007yy, li:2008oi, Sunayama:2015ue}.  However, the degree to which the properties of galaxies themselves are influenced by assembly bias remains an open question.  SHAM algorithms that employ measures of $V_{\rm max}$ (like the RV model) already induce assembly bias into galaxies because concentration, and therefore circular velocity, is correlated with formation history at fixed halo mass \citep{Zentner:2014ki}.  Conditional abundance matching \citep[CAM,][]{Hearin:2014hh} extends the SHAM framework by allowing for two or more halo properties to influence how galaxy properties are assigned.  For example, CAM has been used to study the dependence of the star-formation and halo formation history \citep{Hearin:2013km, Hearin:2014hh, Watson:2015gq, Saito:2015vi, Paranjape:2015uy}.  SHAM models that assign stellar mass and star-formation in a self consistent manner are still in development.

@@

In this section we examine whether introducing assembly bias into the galaxy-halo mapping can increase the satellite fraction and galaxy clustering signal in mass based SHAM models.  Specifically, we consider a model where $M_{*}$ is correlated with the formation history of the (sub-)halo it occupies such that early forming haloes host more massive galaxies than late forming haloes at {\em fixed} peak (sub-)halo mass.  At the same time, we wish to preserve a simple SMHM relation.  One requirement of such model is for there to be a significant amount of scatter in the SMHM relation, $\sigma_{\log(M_*)}$.  The scatter provides a dynamic range in stellar mass at fixed halo mass over which $M_*$ can be correlated with formation history.  While there are many measures of halo formation history, for this work, we employ the redshift a halo reaches a fraction, $\rm f$, of its peak halo mass, $z_{\rm f}$ (see Appendix \ref{sec:halo_properties} for details on how $z_{\rm f}$ is calculated).

We begin by modifying the RM model using the CAM technique.  First, we add log-normal scatter to the assigned stellar masses at fixed $M_{\rm peak}$.  Second, we apply a correlation between $M_{*}$ and $z_{\rm f}$ at fixed $M_{\rm peak}$.  We accomplish this by binning (sub-)haloes in small 0.1 dex $M_{\rm peak}$ bins.  We then rank order sub-haloes by $z_{\rm f}$ and galaxies by $M_{*}$, assigning the most massive galaxies to the earliest forming (sub-)haloes.  We use ${\rm f}=0.5$, the redshift where a (sub-)halo reaches half its peak mass.  Furthermore, parametrize the strength of this correlation by degrading the rank ordering as described in appendix \ref{appendix:ranks} in order to reduce the effect of assembly bias on $M_*$.

%While ideally we would include evolution in the SMHM relation, but evolution introduces significant complexity into the CAM modelling when including assembly bias.  In principle, the stellar mass growth history of each (sub-)halo should be correlated with formation history up to that point, such that the stellar mass function at each redshift is reproduced.  In this way, a truly consistent model for stellar mass growth could be implemented; however, such a model is beyond the scope of this paper.  Instead, we simply examine the degree and phenomena of such a model on the galaxy clustering.  Using a non-evolving SMHM relation will is a logical place to begin.

We find that using a fixed amount of scatter, $\sigma_{\log(M_*)} = 0.18$ dex (similar to the other mass based models), and a constant strong correlation between $M_{*}$ and $z_f$ produces a poor fit to the observed galaxy clustering signal.  In particular, the fraction of massive galaxies that are satellites becomes unrealistically large and the clustering signal increases dramatically for the high stellar masses.  At the same time, there is relatively little effect on the clustering signal at lower stellar masses.  This can be understood as a result of the SMHM relation becoming steep at the low mass end, and $\sim 0.18$ dex scatter in the SMHM relation is not sufficient to increase the satellite fraction enough.

@@

We have shown that there is no published sub-halo abundance matching (SHAM) model that simultaneously:
\begin{enumerate}
\item fits the stellar mass function as a function of redshift, $\phi(M_*,z)$,
\item fits the clustering of galaxies at z=0, $w_p(r_p)$,
\item and uses only identified, extant, sub-haloes in high resolution dark matter simulations.
\end{enumerate}
Models that fit observations of galaxy clustering are incompatible with the observed evolution of the stellar mass function and expectations for the build up of stellar mass in haloes.  Conversely, SHAM models which self-consistently fit the stellar mass function as it evolves significantly under-predict galaxy clustering signals at small scales ($\leq 1 h^{-1}{\rm Mpc}$).  This tension exposes a clustering ``crisis'' for SHAM.

Of the five different SHAM models examined in this work, only SHAM based on rank ordering (sub-)haloes by peak maximum circular velocity, $V_{\rm peak}$, (RV  model) results in a robust galaxy clustering signal consistent with observations.  This finding is in-line with previous work that finds $V_{\rm peak}$ is the best quantity to consider when fitting galaxy clustering observations \citep{Reddick:2013gi, Lehmann:2015vi}.  However, an implicit assumption in $V_{\rm peak}$ SHAM is that the stellar mass-$V_{\rm peak}$ (SMVP) relation does not evolve. Because haloes grow their potential wells early \citep[e.g.][]{Bosch:2014cu}, a non-evolving SMVP relation results in haloes building up stellar mass too early.

>>>>

\title[The Galaxy Clustering Crisis]{The Galaxy Clustering Crisis in Abundance Matching}
\author[Campbell et al.]
{\parbox[t]{\textwidth}{
Duncan Campbell$^{1}$\thanks{E-mail: duncan.campbell@yale.edu},
Frank C. van den Bosch$^{1}$,
Nikhil Padmanabhan$^{2, 3}$, \\ 
Andrew Zentner$^{4}$,
Johannes Lange$^{1}$,
Yao-Yuan Mao$^{4}$, and
Fangzhou Jiang$^{5}$
} \\
\vspace*{3pt} \\
$^{1}$Department of Astronomy, Yale University, New Haven, CT 06511, USA\\
$^{2}$Department of Physics, Yale University, New Haven, CT 06520, USA \\
$^{3}$Yale Center for Astronomy and Astrophysics, Yale University, New Haven, CT 06520, USA\\
$^{4}$Department of Physics and Astronomy \& Pittsburgh Particle Physics, Astrophysics, and Cosmology Center (PITT PACC),\\ University of Pittsburgh, Pittsburgh, PA 15260
$^{5}$Racah Institute of Physics, The Hebrew University, Jerusalem 91904, Israel
}

@@

\begin{abstract}
Galaxy clustering on small scales is significantly under-predicted by sub-halo abundance matching (SHAM) models that populate (sub-)haloes with galaxies based on peak halo mass, $M_{\rm peak}$.  SHAM methods based on the peak maximum circular velocity, $V_{\rm peak}$, have had much better success.  The primary reason $M_{\rm peak}$ based models fail is the relatively low abundance of satellite galaxies produced in these models compared to those based on $V_{\rm peak}$.  Despite success in predicting clustering, simple $V_{\rm peak}$ based SHAM models have their own problems, and there is little theoretical reason to prefer $V_{\rm peak}$ over $M_{\rm peak}$ as a proxy for stellar mass.  We identify three possible remedies that could save mass based models.  1.) SHAM models require a significant population of ``orphan'' galaxies as a result of artificial disruption of sub-haloes in modern dark matter simulations, 2.) satellites must grow significantly after the time of accretion, and/or 3.) stellar mass is significantly affected by halo assembly history.  Regardless of the particulars, we show that popular SHAM models based on $M_{\rm peak}$ cannot be correct as presented.  Either $V_{\rm peak}$ truly is a better predictor of stellar mass, or SHAM models are missing vital component(s) that significantly affect galaxy clustering.
\end{abstract}

@@

SHAM, in its most simple form, assumes that the cumulative abundance of galaxies and haloes can be used to map galaxy properties uniquely onto (sub-)haloes.  Using stellar mass, $M_{*}$, and a halo mass proxy $\mathcal{M}$, SHAM assumes:
%
\begin{linenomath}
\begin{equation}
\label{eq:sham}
N(>M_*) = N(>\mathcal{M}) \implies M_{*} = f(\mathcal{M})
\end{equation} 
\end{linenomath}
%
where $f(\mathcal{M})$ is some monotonically increasing function.  This maps the most massive galaxies to the most massive haloes.  The function $f(\mathcal{M})$ may be determined non-parametrically by matching the rank orders of galaxies and haloes.  We refer to this method of SHAM as ``rank order'' abundance matching.  However, we also consider another class of SHAM models where $f(\mathcal{M})$ is parametrized.  This introduces the possibility that the observed abundance of galaxies is not strictly matched; however, it is generally the goal of such models to preserve this characteristic.  We consider these two types of models similar enough to be contained within the class of models we refer to as SHAM.

@@

\cite{Moster:2013ab} parametrize the SMHM relation as a function of halo mass and redshift to account for evolution.  Furthermore, they assume the stellar mass of satellite galaxies is determined by the mass of the sub-halo at the time of its accretion into a more massive host-halo for the first time.  The functional form is given by:
%
\begin{linenomath}
\begin{equation}
\label{eq:moster_smhm}
\frac{\langle M_{*} | \mathcal{M}\rangle_{\rm med}(a)}{\mathcal{M}} = 2N\left[ \left(\frac{\mathcal{M}}{{M}_1} \right)^{-\beta} + \left( \frac{\mathcal{M}}{{M}_1} \right)^{\gamma} \right]^{-1}
\end{equation}
\end{linenomath}
%
where $\langle\rangle_{\rm med}$ indicates the median.  The evolution of the parameters is given by:
%
\begin{linenomath}
\begin{align}
\label{eq:smhm_params}
\log[{M}_1(a)]& = {M}_{10} + {M}_{11}(1-a) \\
N(a) &= N_{10} + N_{11}(1-a) \\
\beta(a) &= \beta_{10} + \beta_{11}(1-a) \\ 
\gamma(a) &=  \gamma_{10} + \gamma_{11}(1-a)
\end{align}
\end{linenomath}
%
where the scale factor is either the instantaneous scale factor, or the one at the time of accretion:
%
\begin{linenomath}
\begin{equation}
a = 
\begin{cases}
    a      & \quad \text{if host-halo}\\
    a_{\rm acc}  & \quad \text{if sub-halo} \\
  \end{cases}
\end{equation}
\end{linenomath}
%
and $\mathcal{M}$ is the instantaneous mass or the mass at accretion, $M_{\rm acc}$, for sub-haloes.  Scatter in the SMHM relation is modelled as a fixed log-normal with $\sigma \approx 0.18$.

@@

\citet{Behroozi:2013fg} make similar assumptions as M13 but utilize a different parametrization given by:
%
\begin{linenomath}
\begin{equation}
\log[\langle M_*|\mathcal{M} \rangle_{\rm med}(a)] = \log(\epsilon M_1) + f(\log(\mathcal{M}/M_1)) - f(0)
\end{equation}
\end{linenomath}
%
where,
% 
\begin{linenomath}
\begin{equation}
f(x) = -\log(10^{\alpha x}+1) + \delta\frac{\left[ \log(1+\exp[x])\right]^{\gamma}}{1+\exp(10^{-x})}
\end{equation}
\end{linenomath}
%
and where $\mathcal{M} = M_{\rm peak}$.  The parameters evolve with redshift as:
%
\begin{linenomath}
\begin{align}
\nu(a) & = e^{-4a^2}  \\
\log[M_1(a)] & =  M_{10} + \nu\left[ M_{1,a}(a-1) + M_{1,z}(z) \right] \\
\log[\epsilon(a)] & = \epsilon_0 + \nu \left[\epsilon_a(a-1) +\epsilon_z (z) \right] + \\ \nonumber 
& ~~~~ \epsilon_{a,2}(a-1)  \\
\alpha(a) & =  \alpha_{0} + \nu\left[ \alpha_a(a-1)\right] \\
\delta(a) & =   \delta_{0} + \nu\left[ \delta_a(a-1) + \delta_z (z)\right] \\
\gamma(a) &= \gamma_{0} + \nu\left[ \gamma_a(a-1) + \gamma_z (z)\right]
\end{align}
\end{linenomath}
% 
When fitting their model, \cite{Behroozi:2013fg} also allow the scatter in the SMHM relation to vary as a function of redshift.  The variation found is consistent with no variation (constant over cosmic time), so we neglect the small variation in the best fit model and use a constant non-varying scatter.  We have checked that including this variation has no appreciable effect on our conclusions.   As shorthand, we will refer to this model as ``B13''.   B13 constrained the parameters of this model by fitting the stellar mass function at multiple epochs, specific star-formation rates, and the cosmic star-formation history between z=0-8.  The values used in this paper are taken directly from \cite{Behroozi:2013fg} and listed in table 2.

@@

\cite{Yang:2012ew} take a different approach than M13 and B13.  They allow for evolution with redshift in a similar manner, but additionally allow satellites to grow or lose mass after the time of accretion.  The SMHM relation for central galaxies is: 
%
\begin{linenomath}
\begin{equation}
\langle M_{*,\rm cen} |  \mathcal{M} \rangle_{\rm med}(z) = M_{0}\frac{\left( \mathcal{M}/M_1 \right)^{\alpha+\beta}}{\left(1+ \mathcal{M}/M_1 \right)^{\beta}}
\end{equation}  
\end{linenomath}
%
where $\mathcal{M}=M_{180b}$, and the parameters\footnote{We have altered the naming scheme somewhat to be more consistent with the other models.} evolve with redshift as:
%
\begin{linenomath}
\begin{align}
\log[M_{0}(z)] &= M_{01} + \gamma_1 z \\
\log[M_{1}(z)] &= M_{11} + \gamma_2 z \\
\alpha(z) &= \alpha_0 + \gamma_3 z \\
\log[\beta(z)] &= \log(\beta_0) + \gamma_4 z + \gamma_5 z^2
\end{align}
\end{linenomath}
%
The mean stellar mass of satellites is determined by interpolating between the stellar mass a satellite would have had at the time of accretion and the stellar mass of a central galaxy with the same halo mass as the satellite at accretion.  This value is determined by a single parameter, $c$:
%
\begin{linenomath}
\begin{align}
\label{eq:y12sat_model}
\langle M_{*,\rm sat}(z, z_{\rm acc})\rangle =&  (1-c)\langle M_{*,\rm cen} |  \mathcal{M}_{\rm acc} \rangle(z_{\rm acc})~+ \\ \nonumber &  c\langle M_{*,\rm cen}|  \mathcal{M} \rangle(z)
\end{align}
\end{linenomath}
%
where $\mathcal{M}_{\rm acc}$ is the halo mass proxy at the time of accretion for the sub-halo.  The case where $c=0$ corresponds to no growth (or mass loss) since the time of accretion.  The case where $c=1$ corresponds to using the same SMHM relation for centrals and satellites at all redshifts.

@@

Furthermore, each of the evolving models was tuned using observations based on different methods of determining stellar mass.  We apply a simple set of conversions in order to homogenize the stellar masses to a single system as described in Appendix \ref{appendix:stellar_mass_conv}.   We show the stellar mass function for a mock based on each model in Fig. \ref{fig:phi}.  The stellar mass functions in each mock vary by up to $\sim 0.2$ dex from our fiducial stellar mass function used to create the RM and RV mocks.  This is the result of each model being tuned with differing:
\begin{enumerate}
\item stellar mass functions, $\phi(M_*,z)$
\item halo mass definitions/(sub-)halo finder
\item cosmologies
\item simulations which are subject to cosmic variance
\end{enumerate}
While in principle each model could be re-fit using the same stellar mass function(s) and simulation, that is beyond the scope of this paper.  We check that simple alterations to the parameters of each model which bring each into better agreement with the LW09 stellar mass function do not have a significant effect on the clustering signal predictions presented in \S \ref{sec:clustering}.  Given this, and the complexity involved in re-fitting each model, we use the parameters as they are quoted in each paper (and reproduced in table 2, with only minor alterations as discussed in the sections above.

@@

Here we compare the RV model to the other models by examining the relation between stellar mass and halo mass that arises in each model, in the RV model by correlation between $V_{\rm peak}$ and $M_{\rm peak}$, and in the other models by design.  In the top panel of Fig. \ref{fig:SMHM_comparison}, we examine the relation between stellar mass and $M_{\rm peak}$ for individual galaxies in the RV mock.  Beyond the ubiquitous shape of the SMHM relation, the most obvious additional trend is that the RV model predicts that sub-haloes ($z_{\rm acc}>0$) host more massive galaxies than host-haloes of equivalent $M_{\rm peak}$.  This is precisely the expectation given that sub-haloes generally have larger values of $V_{\rm peak}$ than host-haloes as shown in the previous section.

@@

We calculate the sub-halo completeness in the simulation as the ratio of the empirical $f_{\rm sub}$ and the fit using eq. \ref{eq:fsub_fit} as shwon in the bottom panel of Fig. \ref{fig:f_sub}.  We then model the completeness, $C_{\rm sub}$, as a function of halo mass as:
%
\begin{linenomath}
\begin{equation}
\label{eq:sub_completeness}
C_{\rm sub}(M_{\rm peak}) = \frac{C_0}{1.0 + \left(\frac{M_0}{M_{\rm peak}}\right)^{\gamma}}
\end{equation}
\end{linenomath}
%
We find $\log(M_0/h^{-1}M_{\odot})=9.980\pm0.003$ and $\gamma=1.27 \pm 0.02$ provide a good fit, as shown by the red line in the bottom panel in Fig. \ref{fig:f_sub}.  For now, we fix $C_0=1.0$, and we examine the possibility of $C_0<1.0$ at the end of this section.  The implicit assumption here is that when sub-haloes are well resolved, as is the case for massive (sub-haloes) at the time of infall, there should be no missing sub-structure.  Our estimation of the completeness is broadly consistent with the 1000 particle threshold found by \citet{Guo:2013fm} using the Millennium simulation suite.  Furthermore, we find that $C_{\rm sub}(M_{\rm peak})$ is very nearly constant with redshift ($z=0-4$).

@@

The Y12 model model when fit to galaxy clustering in \cite{Yang:2012ew}, which used an analytic halo model and let the sub-halo abundance be a free parameter, finds results which are consistent with our finding here that $C_0 \sim 0.5$, namely that the sub-halo abundance needs to be approximately a factor of two larger than that in Bolshoi.  Furthermore, \cite{Jiang:2016dw} find that the unevolved (peak sub-halo mass) surviving (not disrupted) sub-halo mass function is approximately a factor of two lower than the unevolved sub-halo mas function.  This suggest a plausible source of the missing sub-haloes needed to host orphans could be artificially disrupted sub-haloes.

@@

Assembly bias is a phenomenon observed in $\Lambda$CDM simulations of structure formation that the clustering of haloes is affected by formation history in addition to mass \citep{Gao:2005ds, Wechsler:2006rg, Gao:2007yy, li:2008oi, Sunayama:2015ue}.  However, the degree to which the properties of galaxies themselves are influenced by the assembly history of their halo remains an open question.  SHAM algorithms that employ measures of $V_{\rm max}$ (like the RV model) already induce assembly bias into galaxies because concentration, and therefore circular velocity, is correlated with formation history at fixed halo mass \citep{Zentner:2014ki}.  Conditional abundance matching \citep[CAM,][]{Hearin:2014hh} extends the SHAM framework by allowing for two or more halo properties to influence how galaxy properties are assigned in the SHAM algorithm.  CAM has been used to study the dependence of star-formation rate and halo formation history \citep{Hearin:2013km, Hearin:2014hh, Watson:2015gq, Saito:2015vi, Paranjape:2015uy}; however, SHAM models that assign stellar mass and star-formation in a self consistent manner are still in development.

In this section we examine whether introducing assembly bias into the galaxy-halo mapping can increase the satellite fraction and galaxy clustering signal in mass based SHAM models.  Specifically, we consider a model where $M_{*}$ is correlated with the formation history of the (sub-)halo it occupies such that early forming haloes host more massive galaxies than late forming haloes at {\em fixed} peak (sub-)halo mass.  While many measures of halo formation history have been employed, for this work we use the redshift a halo first achieves a fraction, $\rm f$, of its peak halo mass, $z_{\rm f}$ (see Appendix \ref{sec:halo_properties} for details on how $z_{\rm f}$ is calculated).

@@

We begin by inducing assembly bias in the RM model.  In order for assembly bias to have any effect on $M_*$, there must be a significant amount of scatter in the SMHM relation, $\sigma_{\log(M_*)} >0$.  Scatter provides a dynamic range in $M_*$ at fixed $M_{\rm peak}$ over which $M_*$ can be correlated with formation history.  In Fig. \ref{fig:scatter_model} we shown some constraints on $\sigma_{\log(M_*)}$  from the literature.   Typical values found for $\sigma_{\log(M_*)}$ are between $0.1-0.2$ dex \citep{More:2011il, Reddick:2013gi, Zu:2015vh, Tinker:2016vu}; however, these values are most strongly constrained at high masses, $M_{\rm vir}>10^{12}~h^{-1}M_{\odot}$.  We add scatter to the SMHM in the RM model as described Appendix \ref{sec:rank_scatter}, parametrizing the level of scatter as a function of halo mass, $\sigma_{\log(M_*)}(M_{\rm peak})$.

To induce a correlation between $M_{*}$ and $z_{\rm f}$ we bin (sub-)haloes in small 0.1 dex $M_{\rm peak}$ bins.  We then rank order (sub-)haloes by $z_{\rm f}$ and galaxies by $M_{*}$, re-assigning the most massive galaxies to the earliest forming (sub-)haloes in the bin.  We use ${\rm f}=0.5$, the redshift where a (sub-)halo reaches half its peak mass.  Furthermore, we parametrize the strength of this correlation by the Spearman’s rank order correlation coefficient, $\rho(M_{\rm vir})$.  We weaken the correlation by degrading the rank ordering as described in appendix \ref{appendix:ranks} in order to reduce the effect of assembly bias on $M_*$.

Initially we try a model with fixed scatter, $\sigma_{\log(M_*)} = 0.18$ dex (similar to the other mass based models), and a constant maximum ($\rho=1$) correlation between $M_{*}$ and $z_f$ produces a poor fit to the observed galaxy clustering signal.  The fraction of massive galaxies that are satellites in this model is unrealistically large, and the clustering signal increases dramatically at high stellar masses.  At the same time, there is relatively little effect on the clustering signal at lower stellar masses.  This can be understood as a result of the SMHM relation becoming steep at the low mass end, and $\sim 0.18$ dex scatter in the SMHM relation is not sufficient to increase the satellite fraction enough.

@@

We have shown that there is no published sub-halo abundance matching (SHAM) model that simultaneously:
\begin{enumerate}
\item fits the clustering of galaxies at z=0, $w_p(r_p)$,
\item reproduces the evolution of the stellar mass function, $\phi(M_*,z)$,
\item and uses only identified, extant, sub-haloes in high resolution dark matter simulations.
\end{enumerate}
Models that fit observations of galaxy clustering are incompatible with the observed evolution of the stellar mass function and expectations for the build up of stellar mass in haloes.  Conversely, SHAM models which self-consistently fit the stellar mass function as it evolves significantly under-predict galaxy clustering signals at small scales ($\leq 1~h^{-1}{\rm Mpc}$).  This tension exposes a clustering ``crisis'' for SHAM.

Of the five different models examined in this work, only SHAM based on rank ordering (sub-)haloes by peak maximum circular velocity, $V_{\rm peak}$, (RV  model) results in a robust galaxy clustering signal consistent with observations.  This finding is in-line with previous work that finds $V_{\rm peak}$ is the best quantity to consider when fitting galaxy clustering observations \citep{Reddick:2013gi, Lehmann:2015vi}.  However, an implicit assumption in $V_{\rm peak}$ SHAM is that the stellar mass-$V_{\rm peak}$ (SMVP) relation does not evolve. Because haloes grow their potential wells early \citep[e.g.][]{Bosch:2014cu}, a non-evolving SMVP relation results in haloes building up stellar mass too early.
